---
title: "Preprocessing Pipeline"
author: "Isaac Treves, Lilian Li, Paul Bloom,..."
date: "2025-02-11"
output: html_document
---
# This preprocessing invovlves python
# you can do everything except optional POS tagging without setting up an environment (not in SL manuscript)
# otherwise, you need an environment with python 3.9

# at present, cleaning repeated special characters (see end) to remove more searches and nonsensical messages, but not ellipses. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rwhatsapp)
library(reticulate)
library(cleanNLP)
library(qdapRegex)
library(readr)

#use_condaenv(condaenv = "py39", conda = "/opt/anaconda3/bin/conda")
```



# Load data
Combine all raw key input files - possible problems , if hidden files in directory


```{r data loading,echo=FALSE,message=FALSE}
dir=""
files = list.files(path = dir)


combine_files = function(dir, fname) {
  id = as.numeric(fname)
  crnt_path = paste0(dir,"/",fname)
  data = list.files(path = crnt_path, pattern = "deidentified_hashed_language.csv", recursive = T)
  check<-data %>%
    map_df(~read_csv(paste0(crnt_path,"/",.)))

  check<-check %>%
    mutate(subjectID = id) %>%
    select(subjectID, tm_message_start, tm_message_end, strinput_text, id_app, cat_tz)
}
# watch out for ghost files in directories if you get any errors ,from terminal you can do ls -a

safeFUN = safely(.f = combine_files)
combined = files %>%
  map_df(~safeFUN(dir = dir, fname = .)) %>%
  .$result %>%
  arrange(subjectID, tm_message_start)

# combined<-files%>%map_df(~combine_files(dir=dir,fname=.)) %>% .$result %>% arrange(subjectID,tm_message_start)
head(combined)
#rm(dir,files)
```
#Optional checking of language sufficiency (>15 days of data) and demographics 
```{r,echo=FALSE}

load("demos.Rda") # not supplied 
combinedcheck<-combined %>% mutate(day= format(tm_message_start, "%Y-%m-%d"))
combinedcheck<-combinedcheck %>% group_by(subjectID) %>% summarize(days=length(unique(day)),message=length(unique(tm_message_start)))
combinedcheck$WeeksLanguageSufficient=combinedcheck$days>15
selfreport<-merge(selfreport,combinedcheck,by.x="ID",by.y="subjectID",all.x=TRUE)
rm(selfreport_itemwise, selfreport_long)
```

```{r,echo=FALSE}
table(selfreport$sex,selfreport$WeeksLanguageSufficient,useNA = "ifany")
table(selfreport$GROUP,selfreport$WeeksLanguageSufficient,useNA = "ifany")
#t.test(selfreport$age~selfreport$WeeksLanguageSufficient)
#t.test(selfreport$age~selfreport$WeeksLanguageSufficient)

```
# Check for null app labels 
mostly a problem for 1100s, so we should make sure not to filter based on social or not
``` {r,echo=FALSE} 

sum(combined$id_app == "<null>")/length(combined$id_app)*100  
check = combined %>%
  group_by(subjectID) %>%
  summarise(n = n(), withoutApp = sum(id_app == "<null>")/length(id_app)*100) %>%
  ungroup() 
```

# Check if additional apps and label with social media
```{r} 
socialMedia = read_csv("apps_list_withai.csv")
#appsCleaned = readxl::read_excel("applications_check.xlsx")
appsCurrentData = combined %>% distinct(id_app) %>% filter(id_app != "<null>") 

appsMore<-appsCurrentData%>%filter(!(appsCurrentData$id_app %in% intersect(appsCurrentData$id_app,socialMedia$ID)))

#appsSocialMedia=socialMedia$ID[socialMedia$is_social_media==1];
#combined<-combined %>% mutate(socialOrNot=ifelse(id_app %in% appsSocialMedia,1,0))
combined <- merge(combined,socialMedia %>% select(ID,Name,Social.or.Nonsocial,is_social_media),by.x="id_app", by.y="ID",all.x=T)
combined<-combined %>% mutate(socialOrNot=ifelse(is_social_media==1,1,0)) %>% select(-is_social_media) %>% rename(app_name=Name) 
rm(socialMedia)

#write_csv(appsMore, "appsMore.csv")
```
remove lock screen
```{r}
combined <- combined %>% filter(id_app != "com.htc.lockscreen" & id_app != "com.android.systemui")
combined <- combined %>% filter(app_name != "Dialer" & app_name != "OnePlus Dialer" & app_name != "Contacts")



```

# beginning cleaning
changed the loading of groovy and pls 
``` {r} 
combined_save=combined # for putting aside combined in case you want to adjust preprocessing
device = read_csv("device list") # get device list , not provided


android = device %>% filter(OS != "iOS") 

# remove empty rows and unreadable characters "\ufffc"
combined = combined %>%
  arrange(subjectID, tm_message_start) %>%
  mutate(strinput_text = gsub("\ufffc","",strinput_text)) %>%
  filter(!(is.na(strinput_text) | strinput_text == "")) 


emojis = emojis 

# ;-; crying face



tmp = combined %>% filter(nchar(strinput_text) == 1) %>% left_join(emojis, by = c("strinput_text" = "emoji")) # remove everything except for emoji

# remove strings with only 1 character other than emojis
combined = combined %>%
  filter(nchar(strinput_text) > 1 | strinput_text %in% emojis$emoji)



tmp = combined %>%
  count(strinput_text) %>%
  ungroup() %>%
  filter(n > 1) %>%
  arrange(-n)

tmp_android = combined %>%
  filter(subjectID %in% android$subject_id) %>%
  count(strinput_text) %>%
  # filter(n > 1) %>%
  arrange(-n)

tmp_ios = combined %>%
  filter(!(subjectID %in% android$subject_id)) %>%
  count(strinput_text) %>%
  filter(n > 1) %>%
  arrange(-n)

# this is a vector of 'automatic' strings: 'Send a chat' to remove, based on Lilian's perusal
# we're looking for exact matches, which is probably not the best
# ideally you would use regular expressions e.g. 'say something in .* chat'
load("android_strings.RData")
sys.txt=test
rm(test)

discord = combined %>% filter(id_app == "com.hammerandchisel.discord"| id_app == "com.hammerandchisel.discord.Share" | id_app == "com.discord" | id_app=="Discord") %>% count(strinput_text) %>% arrange(-n)


plsCom=readLines("pls command.txt")
plsCom=data.frame(command=plsCom[!plsCom==""])
pls = "^pls (animals"
for (k in 2:dim(plsCom)[1]) {
  pls = paste(pls, plsCom$command[k], sep = "|")
}
pls = paste(pls, ").*", sep = "")


groovyCom=readLines("groovy command.txt")
groovyCom=data.frame(command=groovyCom[!groovyCom==""])
groovy = "^/(play"
for (k in 2:dim(groovyCom)[1]) {
  groovy = paste(groovy, groovyCom$command[k], sep = "|")
}
groovy = paste(groovy, ").*", sep = "")


# the reason this doesn't replace valid text is it has to start with pls or groovy to be replaced
check = discord %>%
  mutate(strinput_text = gsub("^owo .*","", strinput_text, ignore.case = T)) %>%
  filter(strinput_text != "Owo" & strinput_text != "owo") %>%
  mutate(strinput_text = gsub("^(Owocdaily|Owocookie|Owocpropose|Owozoo).*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub(pls,"", strinput_text, ignore.case = T)) %>%
  filter(strinput_text != "Plsb" & strinput_text != "Plsnm" & strinput_text != "Ppls bal") %>%
  mutate(strinput_text = gsub(groovy,"", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^/(mx|ma|mg|hx|ha|hg|wx|wa|wg|im|ima|tu|kakera|overview|settings|help|commandsearch|pokeslot).*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^\\$[a-z].*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^:-[a-z].*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^![a-z].*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^p![a-z].*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^\\?[a-ln-z].*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^\\?mute .*","", strinput_text, ignore.case = T)) %>%
  filter(!(is.na(strinput_text) | strinput_text == ""))

reddit = combined %>% filter(id_app == "com.reddit.Reddit"| id_app == "com.reddit.frontpage" | id_app == "com.christianselig.Apollo" | id_app == "free.reddit.news" | id_app=="Reddit") %>% count(strinput_text) %>% arrange(-n) #

# looking at rapid messages  "where are", "where are you"
check = combined %>%
  group_by(subjectID) %>%
  mutate(timeDiff = difftime(tm_message_start, lag(tm_message_start), units = "secs"),
         val_contain = stringr::str_detect(tolower(strinput_text), coll(tolower(lag(strinput_text))))) %>%
  mutate(sus = ifelse(lead(val_contain == T) & id_app == lead(id_app), 1, 0)) %>%
  ungroup() %>%
  filter(!(is.na(strinput_text) | strinput_text == ""))

tmp = check %>%
  filter(lag(sus) == 1) 

quantile(tmp$timeDiff, .95) #95% pctile is 49 sec

```
# Full cleaning 


``` {r, echo= FALSE}

# remove empty rows and unreadable characters "\ufffc"
combined = combined %>%
  arrange(subjectID, tm_message_start) %>%
  mutate(strinput_text = gsub("\ufffc","",strinput_text)) %>%
  filter(!(is.na(strinput_text) | strinput_text == "")) 

#  for android only, clean 
combined = combined %>%
  mutate(strinput_text = ifelse(subjectID %in% android$subject_id, gsub(".*…$","",strinput_text), strinput_text)) %>%
  mutate(strinput_text = ifelse(subjectID %in% android$subject_id, gsub("^Search.*","",strinput_text), strinput_text)) %>%
  mutate(strinput_text = ifelse(subjectID %in% android$subject_id, gsub("^Reply to.*","",strinput_text), strinput_text)) %>%
  mutate(strinput_text = ifelse(subjectID %in% android$subject_id & strinput_text %in% sys.txt, "", strinput_text)) %>%
  mutate(strinput_text = gsub("Get Outlook for Android","",strinput_text)) %>%
  mutate(strinput_text = gsub("Add A Heading","",strinput_text)) %>%
  mutate(strinput_text = ifelse(subjectID %in% android$subject_id, gsub("^\u2022[^ ].*","",strinput_text, ignore.case = T), strinput_text)) %>% #remove texts starting bullets
  mutate(strinput_text = ifelse(subjectID %in% android$subject_id, gsub(".*[^ ]\u2022$","",strinput_text, ignore.case = T), strinput_text)) %>% #remove texts ending bullets
  mutate(strinput_text = ifelse(subjectID %in% android$subject_id, gsub("#","3",strinput_text), strinput_text)) %>%
  filter(!(is.na(strinput_text) | strinput_text == ""))

# now we're replacing - not just to discord but the strings are probably rare
combined = combined %>%
  mutate(strinput_text=gsub("say something in .* chat","",strinput_text,ignore.case=T)) %>%
  mutate(strinput_text=gsub("^Send a chat .*","",strinput_text,ignore.case=T)) %>%
  mutate(strinput_text=gsub("^Add Reply.*","",strinput_text,ignore.case=T)) %>%
  mutate(strinput_text=gsub("Send .* message .*","",strinput_text,ignore.case=T)) %>%
  mutate(strinput_text=gsub("Type your message here...","",strinput_text,ignore.case=T)) %>%
  mutate(strinput_text=gsub("New message","",strinput_text,ignore.case=T)) %>%
  mutate(strinput_text = gsub("^owo .*","", strinput_text, ignore.case = T)) %>%
  filter(strinput_text != "Owo" & strinput_text != "owo") %>%
  mutate(strinput_text = gsub("^(Owocdaily|Owocookie|Owocpropose|Owozoo).*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub(pls,"", strinput_text, ignore.case = T)) %>%
  filter(strinput_text != "Plsb" & strinput_text != "Plsnm" & strinput_text != "Ppls bal") %>%
  mutate(strinput_text = gsub(groovy,"", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^/(mx|ma|mg|hx|ha|hg|wx|wa|wg|im|ima|tu|kakera|overview|settings|help|commandsearch|pokeslot).*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^\\$[a-z].*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^:-[a-z].*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^![a-z].*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^p![a-z].*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^\\?[a-ln-z].*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^\\?mute .*","", strinput_text, ignore.case = T)) %>%
  mutate(strinput_text = gsub("^r/.*","", strinput_text, ignore.case = T)) %>%
  filter(!(is.na(strinput_text) | strinput_text == ""))

# removing fast responses 
combined = combined %>%
  group_by(subjectID) %>%
  mutate(timeDiff = difftime(tm_message_start, lag(tm_message_start), units = "secs"),
         val_contain = stringr::str_detect(tolower(strinput_text), coll(tolower(lag(strinput_text))))) %>%
  filter(!(lead(val_contain == T) & id_app == lead(id_app) & lead(timeDiff) <= 60)) %>%
  ungroup() %>%
  filter(!(is.na(strinput_text) | strinput_text == ""))


# removing twitter urls and mentions 
combined = combined %>%
  mutate(text_clean = rm_url(strinput_text, pattern = pastex("@rm_twitter_url", "@rm_url"), extract = F)) %>% #remove URLs
  mutate(text_clean = gsub('\\S*@\\S*\\s?', '', text_clean)) %>% #remove mentions and emails
  mutate(text_clean = gsub(' +',' ', text_clean)) %>% #remove extra whitespaces
  mutate(text_clean = gsub("^[[:space:]]*","", text_clean)) %>% #remove leading space
  mutate(text_clean = gsub("[[:space:]]*$","", text_clean)) %>% #remove trailing space
  filter(!(is.na(text_clean) | text_clean == ""))
```

# function for removing search strings , used later
```{r, echo=FALSE}
remove_search_text_entries <- function(text, 
                                            special_char_threshold = 3) {
  # Conditions
  starts_with_q <- grepl("^q=", text, ignore.case = TRUE)
  contains_chrome <- grepl("chrome-mobile", text, ignore.case = TRUE)
  
  # Count special characters using regex
  special_counts <- lengths(regmatches(text, gregexpr("[=_&+%#@$^*(){}\\[\\]|\\\\:;\"'<>,./`~]", text)))
  too_many_specials <- special_counts > special_char_threshold
  
  url_like <- grepl("[=_&+%]{3,}", text)
  base64_like <- grepl("[A-Za-z0-9]{50,}", text)
  query_like <- grepl("(&[a-zA-Z]+=|\\?[a-zA-Z]+=)", text)
  
  # Combine all conditions
  should_remove <- starts_with_q | contains_chrome | 
                   too_many_specials | url_like | base64_like | query_like
  
  return(should_remove)
}

```

# doing some python in R, this takes cares of emails and ursl etc. will take time to build all packages , this isn't simple regular expressions
# + final removal of search queries that aren't explictly urls , needs to load function in other cell 'remove_search_text_entries", this part is slower
```{r call python script}


library(reticulate)

#py_install("ekphrasis")

py_install("ekphrasis",pip=TRUE)
# Run the Python script
combined_copy=combined # just to make sure this script doesn't do anything weird
rm(py)
py$strinput_text <- r_to_py(combined_copy$text_clean)

py_run_string("
from ekphrasis.classes.preprocessor import TextPreProcessor

# doesn't work as well as regular expressions for .com 
text_processor = TextPreProcessor(
    normalize=['url', 'email', 'percent', 'money', 'phone', 'user', 'time', 'date'],
    annotate={},
    segmenter='twitter',
    corrector='twitter',
    unpack_hashtags=True,
    unpack_contractions=False)
    #spell_correct_elong=True # I think we don't want to replace elongated words, but # this doesn't work super well

def remove_tags(doc):
    doc = ' '.join(word for word in doc.split() if not (word[0]=='<' and word[-1] == '>'))
    return doc
    
#text = 'Check out this URL: https://example.com, email me at user@example.com, and #call me at 1234567890! #awesome'

# Process each string in the R dataframe column
processed_texts = []
for text in strinput_text:
    processed_text = text_processor.pre_process_doc(text)
    processed_text=remove_tags(processed_text)
    processed_texts.append(processed_text)
    
# Store the processed texts back in R
r.processed_texts = processed_texts
")
# Access the processed texts in R

combined_copy$text_clean <- processed_texts
combined_copy<-combined_copy%>%filter(!(is.na(text_clean) | text_clean == ""))
combined_copy$toremove<-remove_search_text_entries(combined_copy$text_clean) # uses function
combined_copy2<-combined_copy%>%filter(toremove==FALSE)

```
# saving batch
```{r,echo=FALSE}
# Define the finalmessages dataframe
finalmessages <- combined_copy2

# Define the base path

path<-""
# Suggested formats - annotate type of preprocessing and date
date=format(Sys.Date(),format = "%Y_%m_%d")
# Construct the file paths
rdata_file <- file.path(path, paste0("", date, ".RData"))
pickle_file <- file.path(path, paste0("", date, ".pkl"))

# Save the R dataframe to an .RData file
save(finalmessages, file = rdata_file)

# Save to a csv
write_csv(finalmessages, "/", date,".csv")  # readr uses UTF-8 by default


#summary <- finalmessages %>% mutate(text_clean=str_to_lower(text_clean)) %>% summarize(n=NROW(text_clean),.by=text_clean) %>% filter(n>50)

# Convert the R dataframe to a Pandas dataframe - if you have unicode problems go down to problem identification
# 

py_install("pandas")
pd <- import("pandas")  # Import the Pandas module
py_df <- pd$DataFrame(finalmessages)  # Convert R dataframe to Pandas dataframe

# Save the Pandas dataframe to a pickle file
py_save_object(py_df, pickle_file)

# Print confirmation
cat("RData file saved to:", rdata_file, "\n")
cat("Pickle file saved to:", pickle_file, "\n")
```

# saving individuals 
```{r, echo=FALSE}
library(readr) # for encoding
finalmessages <- combined_copy2
date=format(Sys.Date(),format = "%Y_%m_%d")
for (i in unique(finalmessages$subjectID)){
  dir=""
  write_csv(finalmessages[finalmessages$subjectID==i,],file=paste0(dir,i,"/Preprocessed_" ,date,".csv"))
}

```





# POS TAGGING (OPTIONAL NOT IN SL MANUSCRIPT)
this sometimes doesn't work depending on python environment., can't find transformer model, not sure why/ works in python. 
let's use the small model  instead - probably too strict for transformers but you can check it out , maybe good for word-level processing of messages 
```{r optional parts of speech tagging, echo=FALSE}
library(cleanNLP)

py_install("cleannlp", pip = TRUE)
py_install("spacy", pip = TRUE)
cnlp_download_spacy("en_core_web_sm")
py_config()  # This will show which Python version/environment is being used


data_6mo = finalmessages %>% mutate(doc_id = row_number())


#cnlp_download_spacy("en_core_web_sm")
cnlp_init_spacy(model_name = "en_core_web_sm") # English, (so ideally do after translate)
tmp_spacy = cnlp_annotate(data_6mo, 
                          text_name = "text_clean", # did not remove punctuations etc. for improved accuracy
                          doc_name = "doc_id",
                          verbose = 100) 
# saveRDS(tmp_spacy, "tmp_spacy.RDS")
```


# OPTIONAL 

# IF YOU HAVE ENCODING PROBLEMS WITH EMOJIS
```{r}
# 1. Create a function to identify problematic entries
# Fast vectorized approach using base R functions
fast_find_problems <- function(df) {
  problem_rows <- c()
  
  for (col in names(df)) {
    if (is.character(df[[col]])) {
      cat("Checking column:", col, "\n")
      
      col_data <- df[[col]]
      
      # Convert, handling NAs properly
      converted <- iconv(col_data, from = "UTF-8", to = "UTF-8")
      
      # Only compare non-NA values
      # Problem if: conversion failed (became NA when original wasn't) OR content changed
      actual_problems <- which(
        # Conversion failed: original was not NA but converted became NA
        (!is.na(col_data) & is.na(converted)) |
        # Content changed: both are not NA but different
        (!is.na(col_data) & !is.na(converted) & col_data != converted)
      )
      
      if (length(actual_problems) > 0) {
        cat("Found", length(actual_problems), "actual encoding problems in column", col, "\n")
        problem_rows <- unique(c(problem_rows, actual_problems))
      }
    }
  }
  
  return(problem_rows)
}
# Run the fast detection
problem_rows <- fast_find_problems(finalmessages)
cat("Total problematic rows:", length(problem_rows), "\n")

```
# APP NAME PROBLEMS? 
```{r}


# Clean only the problematic entries in app_name
for (i in problem_rows) {
  if (!is.na(finalmessages$app_name[i])) {
    tryCatch({
      # Try multiple cleaning approaches
      cleaned <- iconv(finalmessages$app_name[i], from = "UTF-8", to = "UTF-8", sub = "")
      
      if (is.na(cleaned)) {
        # More aggressive - remove all non-ASCII
        cleaned <- gsub("[^\x01-\x7F]", "", finalmessages$app_name[i])
      }
      
      if (is.na(cleaned) || nchar(cleaned, allowNA = TRUE) == 0) {
        # Nuclear option - keep only alphanumeric and basic punctuation
        cleaned <- gsub("[^A-Za-z0-9 .,!?()-]", "", finalmessages$app_name[i])
      }
      
      finalmessages$app_name[i] <- cleaned
      
    }, error = function(e) {
      # Last resort
      finalmessages$app_name[i] <- "[CORRUPTED]"
    })
  }
}

cat("Cleaned", length(problem_rows), "problematic app_name entries\n")

```
